---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# COMET

<!-- badges: start -->
<!-- badges: end -->

The goal of COMET is to simulate provide an open source framework to simulate the United States Lung Allocation System

## Installation

You can install the development version of COMET like so:
```{r, include=FALSE}
library(COMET)
```


```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("ClevelandClinicQHS/COMET")
```

## Basic example
This will simulate the 2015 Lung Allocation Score (LAS) policy for 400 days (`r1`) and then the same scenario under the Composite Allocaiton Score (CAS) policy (`r2`).
```{r example}
library(COMET)

r1 <- run_simulation(days = 400, can_start = 1000,
                     match_alg = match_las, wl_model = "LAS15", post_tx_model = "LAS15",
                     wl_weight = 2, post_tx_weight = 1, wl_cap = 365, post_tx_cap = 365, seed = 26638)

r2 <- run_simulation(days = 400, can_start = 1000,
                     match_alg = match_cas, wl_model = "CAS23", post_tx_model = "CAS23",
                     wl_weight = 0.25, post_tx_weight = 0.25, wl_cap = 365, post_tx_cap = 1826,
                     bio_weight = .15, pld_weight = 0.05, peds_weight = 0.2, efficiency_weight = 0.1, seed = 26638)

```

A simple way to evaluate the COMET runs, one can also look at the statistics by diagnosis group or another variable. For details on what is reported run `help('eval_simulation')`.
```{r comparison of simualtions}
# see how many were transplanted, donor count, waitlist deaths
eval_simulation(r1)

eval_simulation(r2)

## See results by diagnosis group
eval_simulation(r1, group = dx_grp)
eval_simulation(r2, group = dx_grp)
```

One can also mix and match the models and weights under the LAS/CAS policies. `r3` simulates the but the LAS policy but with the CAS waitlist and post-transplant mortality models (post-transplant survival truncated at 1 year). `r4` is the same scenario as `r3`, but waitlist mortality receives the same weight as post-transplant (1 to 1).
```{r example2}
r3 <- run_simulation(days = 400, can_start = 1000,
                     match_alg = match_las, wl_model = "CAS23", post_tx_model = "CAS23",
                     wl_weight = 2, post_tx_weight = 1, wl_cap = 365, post_tx_cap = 365, seed = 26638)

r4 <- run_simulation(days = 400, can_start = 1000,
                     match_alg = match_las, wl_model = "CAS23", post_tx_model = "CAS23",
                     wl_weight = 1, post_tx_weight = 1, wl_cap = 365, post_tx_cap = 365, seed = 26638)
```

## A more indepth demonstration to show what is going on behind the scenes
### Generates candidates and donors
One major advantage of the simulation is the use of synthetic candidates and donors.
```{r generation}
set.seed(26638)
syn_cands <- gen_and_spawn_candidates(days = 10)
syn_dons <- gen_and_spawn_donors(days = 10)
head(syn_cands)
head(syn_dons)
```

For this scenario I'm going to add `r max(r2$all_candidates$listing_day)` days to the listing and recovery day to `syn_cands` and `syn_donors` and also make sure the candidate and donor ids do not conflict. I'm going to continue the results from CAS simulation (`r2`). The notation for the LAS simulation (`r1`) is very similar.
```{r iterate}
syn_cands <- dplyr::mutate(syn_cands,
                           listing_day = listing_day + max(r2$all_candidates$listing_day),
                           c_id = c_id + max(r2$all_candidates$c_id))

syn_dons <- dplyr::mutate(syn_dons,
                          recovery_day = recovery_day + max(r2$all_candidates$listing_day),
                          d_id = d_id + max(r2$all_donors$d_id))

r2cc <- dplyr::mutate(r2$current_candidates,
                      days_on_waitlist = 401 - listing_day)

r2rd <- dplyr::mutate(r2$recipient_database,
                      days_after_tx = 401 - transplant_day)

l <- list("current_candidates" = r2cc,
          "recipient_database" = r2$recipient_database,
          "waitlist_death_database" = r2$waitlist_death_database,
          "post_tx_death_database" = r2$post_tx_death_database,
          "non_used_donors" = r2$non_used_donors)

l_1 <- iteration(401, syn_cands, syn_dons, include_matches = FALSE, updated_list = l, match_alg = match_cas,
                 wl_model = "CAS23", post_tx_model = "CAS23", wl_weight = 0.25, post_tx_weight = 0.25,
                 wl_cap = 365, post_tx_cap = 1826, bio_weight = .15, pld_weight = 0.05,
                 peds_weight = 0.2, efficiency_weight = 0.1)

l_2 <- iteration(402, syn_cands, syn_dons, include_matches = FALSE, updated_list = l_1, match_alg = match_cas,
                 wl_model = "CAS23", post_tx_model = "CAS23", wl_weight = 0.25, post_tx_weight = 0.25,
                 wl_cap = 365, post_tx_cap = 1826, bio_weight = .15, pld_weight = 0.05,
                 peds_weight = 0.2, efficiency_weight = 0.1)

nrow(l_2$waitlist_death_database) - nrow(l_1$waitlist_death_database)
nrow(l_2$post_tx_death_database)-nrow(l_1$post_tx_death_database)
```

Looking in depth, `r nrow(l_2$waitlist_death_database)-nrow(l_1$waitlist_death_database)` died on the waiting list, and `r nrow(l_2$post_tx_death_database)-nrow(l_1$post_tx_death_database)` died post transplant on Day 402 in this scenario. The following would run ten days sequentially
```{r full_iterate}
l_i <- l
for(i in 401:410){
  l_i <- iteration(i,
                   syn_cands, syn_dons, include_matches = FALSE, updated_list = l_i,
                   match_alg = match_cas,  wl_model = "CAS23", post_tx_model = "CAS23",
                   wl_weight = 0.25, post_tx_weight = 0.25, wl_cap = 365, post_tx_cap = 1826,
                   bio_weight = .15, pld_weight = 0.05, peds_weight = 0.2, efficiency_weight = 0.1)
  
}
```

### Within each day
#### Updating candidates/recipients
`update_patients` is one of the steps that occur each day, it will output a list of those that died, were removed for other reasons if on the waiting list, those that are still alive and updated characteristics. Currently, the only characteristic that is updated is the candidate's/recipient's age, but it is made to be flexible to adjust any clinical characteristic (i.e. $\textrm{pCO}_2$,$\textrm{FEV}_1$, bilirubin, etc.)

Within `update_patients` are 

* `identify_deaths` returns a dataset of those who died either on the waiting list or post-transplant^[If one wonders what model is "CAS23r", it is the baseline hazard for the "CAS23" models 'recalibrated' after research and initial simulations runs showed that the "CAS23" model did not accurately simulate waitlist and post-transplant deaths.]

* `identify_removals` returns a dataset of those who were removed on the waiting list^[Models for waitlist removal for other reasons are not part of the LAS/CAS and were modeled separately. The maximum day for this 2365 days after being placed on the waiting list]

```{r updating}
update_patients(patients = r2cc, model = "CAS23r", 
                elapsed_time = days_on_waitlist, pre_tx = TRUE, cap = 365, date = 401)

identify_deaths(patients = r2cc, model = "CAS23r",
                elapsed_time = days_on_waitlist, pre_tx = TRUE, cap = 365, date = 401)

identify_removals(patients = r2cc, elapsed_time = days_on_waitlist, cap = 2365)

## post-transplant deaths
update_patients(r2rd, model = "CAS23r", 
                elapsed_time = days_after_tx, pre_tx = FALSE, cap = 1826, date = 401)

```

#### Matching
`match_cas` and `match_las` are the two main functions that handle screening, matching and acceptance.
```{r matching}
##
dons_401 <- dplyr::filter(syn_dons, recovery_day == 401)

mtc <- match_cas(cands = r2cc, dons = dons_401, wl_model = "CAS23", post_tx_model = "CAS23", wl_weight = 0.25,
                 post_tx_weight = 0.25, wl_cap = 365, post_tx_cap = 1826, bio_weight = .15,
                 pld_weight = 0.05, peds_weight = 0.2, efficiency_weight = 0.1)

mtc

```

#### Screening
These functions are designed to input candidates and donors respectively. Here are displayed screening for height (`hm1`), blood type (`am1`), count (`cm1` double vs single transplant), and a distance calculation (`lm1`). One can also create addition functions for screening
```{r screening}
## within matching
## there is screening
## Screening is done for height, blood type, organ count,
hm1 <- height_screen(cands = r2cc, dons = dons_401)
hm1
## returns if compabitible for single, double transplant, if a double transplant they will match for double lung
am1 <- abo_screen(cands = r2cc, dons = dons_401)
am1
## returns abo_exact abo compatible organs since that is a tie breaker in the LAS

## Ensure double with double, singles may take a double if only is assigned or either can
cm1 <- count_screen(cands = r2cc, dons = dons_401)
cm1

## Distance calculation 
lm1 <- dist_calc(cands = r2cc, dons = dons_401)
lm1


```

In our simulation you will see to rankings `ov_rank` which stands for overall ranking and `offer_rank`. `ov_rank` is the candidates rank with respective to all candidates on that day based on their LAS/CAS score, regardless of donor/offer. `offer_rank` is the rank of the candidate with respect to all compatible candidates to the given donor. The part of the CAS score that does not change called the "sub-CAS" calculates the `ov_rank`

```{r over rank}
## ranking
lr1 <- calculate_sub_cas(r2cc, wl_model = "CAS23", post_tx_model = "CAS23", wl_weight = 0.25,
                         post_tx_weight = 0.25, wl_cap = 365, post_tx_cap = 1826, bio_weight = .15,
                         pld_weight = 0.05, peds_weight = 0.2)

lr1
```

#### Offering and Acceptance
Once all of the screening criteria, the candidates are matched on all screening criteria and made the offers in successive order based on either the CAS or LAS policy. There is an `acceptance_prop` function which will calculate the acceptance probability and run binomial acceptance for whether or not the candidate will accept the organ. And lastly `transplant_candidates` finds the first candidate that accepts the donor lung and transplants them so the donor identifier `d_id` is tracked with the candidate.
```{r offers}

## matching and offering rank
## Pra screening is not currently implemented,
## you'll notice that if you run the function it returns a combination of every candidate and donor
offs <- cas_offer_rank(hm1, am1, cm1, lm1, overall_ranking = lr1, efficiency_weight = 0.1)
offs

set.seed(26638)
acceptance_prob(offs, dons = dons_401, cands = r2cc)

transplant_candidates(matches = mtc)
```

## Other tips and tricks

You will notice a `pra_screen` function. This is a place-holder for now that returns a grid of every candidate and donor pairing. Future implementation will incorporate PRA screening criteria.
```{r pra}
pra_screen(cands = r2cc, dons = dons_401)
```

If you want to return the matches along with the simulated output. All one has to do in the `run_simulation` or `iteration` function is set `include_matches = TRUE`. The output can be seen below is a nested tibble with a vector of donor ids (`d_id`) that are ranked in order for the subsequent offer, called `all_matches`
```{r in_match}
l_1m <- iteration(401, syn_cands, syn_dons, include_matches = TRUE, updated_list = l, match_alg = match_cas,
                 wl_model = "CAS23", post_tx_model = "CAS23", wl_weight = 0.25, post_tx_weight = 0.25,
                 wl_cap = 365, post_tx_cap = 1826, bio_weight = .15, pld_weight = 0.05,
                 peds_weight = 0.2, efficiency_weight = 0.1)

l_1m$all_matches
```

## A few helpful tips
`eval_simulation` behind the seens runs a function callled `concat2`. It concatenates the several candidates/recipients datasests to one and calculates survival data as well. By default it will use the minimum and maximum number of days simulated from `all_candidates` in a COMET object, but those can also be specified if one wanted to exclude candidates after day 30 `min_enroll_date` and before day 300 days `max_enroll_date` or censor post-transplant early at day 330 `post_tx_censor_date`. Waitlist censoring `wl_censor_date` is also possible to change, but we would recommend keeping it the same as the `max_enroll_date`. 
```{r concat}
concat2(r1)
concat2(r1, min_enroll_date = 31, max_enroll_date = 300, wl_censor_date = 300, post_tx_censor_date = 330)

```

The function `spec_day` will allow one to grab the waiting list on a specific day.
```{r spec}
spec_day(r1, day = 300)
```

`rmst` will output the restricted mean survival time (RMST) or the Area Under the Curve (AUC) for the LAS and CAS waitlist mortality models as well as the linear predictor. The linear predictor is calculated using the `calc_post_tx_cas23` and `calc_wl_cas23` and  other respective functions for the waitlist and post-transplant models for the CAS/LAS policies. Currently implemented are the 2015 LAS (`"LAS15"`), 2021 LAS (`"LAS21"`) and 2023 CAS (`"CAS23"`). This also runs `expected_survival` which actually calculates the AUC for the candidate based on their linear predictor.

```{r rmst}
rmst("LAS21", cand_data = r2cc, cap = 365, wl = TRUE)
rmst("CAS23", cand_data = r2cc, cap = 1826, wl = FALSE)
```

One can calculate RMST if one supplies a vector of survival times, or one can calculate the survival given the raw survival value on a given day. For example below we will use Day 364 of the CAS23 waitlist survival model which has a baseline survival value of 0.994390.
```{r exp_surv}
expected_survival(lp = c(0, 1, -0.5), 0.994390)
```

